---
title: "Seurat_test BDS3"
output: rmdformats::readthedown
date: "`r Sys.Date()`"
---

Активізувати вікно коду - Ctrl+ALT+I
=
# Це заголовок
## Це підзаголовок
### Це під під заголовок 

###### і так далі 

## Встановлення пакетів

```{r}
# install.packages('Seurat')
# 
# remotes::install_github("satijalab/seurat-data", "seurat5", quiet = TRUE)
# 
# BiocManager::install('Seurat')

```

```{r}

library(dplyr)
library(Seurat)
library(patchwork)

# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "/Users/korshe/bds3_intro_to_r/filtered_gene_bc_matrices/hg19/")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
pbmc

```


```{r}
pbmc.data[c("C1orf185", "TCL1A", "MS4A1",'CDKN2C'), 14:26]
```


# QC and selecting cells for further analysis 
# Якість контролю даних 

```{r}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")

pbmc[["percent.A"]] <- PercentageFeatureSet(pbmc, pattern = "^A")

Agenes <- pbmc[["percent.A"]]

# pbmc[["percent.mt"]]


mitoch <- pbmc[["percent.mt"]]
mitoch$percent.mt

mitoch$sum <- mitoch$percent.mt + 10
hist(mitoch$percent.mt)

```


```{r}
head(pbmc@meta.data, 5)
meta_data <- pbmc@meta.data

pbmc@assays$RNA

```


```{r}
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0.01 )

# VlnPlot(pbmc, features = c( "percent.mt", "nFeature_RNA", "nCount_RNA"), ncol = 2)
# 
# VlnPlot(pbmc, features = c( "percent.mt"), ncol = 2)

```

1 - мітохондрії
2 - унікальні гени
3 - кількість підрахунків цих генів

```{r}
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r}

pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5 )

# pbmc

pbmc2 <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5 & nCount_RNA < 5000 & nCount_RNA > 1000 )


```

& - і 
| - або [[pipe]]

# Normalizing the data | нНомралізація даних

```{r}
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

pbmc <- SCTransform(pbmc)

```

```{r}
# pbmc@assays$SCT@data

sct_mtrx <- as.matrix(pbmc@assays$SCT@data)
raw_mtrx <- as.matrix(pbmc@assays$SCT@counts)

raw_mtrx$mean_value <- rowMeans(as.matrix(raw_mtrx))

```

# Identification of highly variable features (feature selection)


```{r}
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
#plot1 + plot2
```

# Perform linear dimensional reduction
```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
print(pbmc[["pca"]], dims = 1:5, nfeatures = 15)

VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
DimPlot(pbmc, reduction = "pca")
DimHeatmap(pbmc, dims = 1, cells = 500, balanced = TRUE)
ElbowPlot(pbmc)

```

```{r}
set.seed(2)
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
head(Idents(pbmc), 5)
```

# Run non-linear dimensional reduction (UMAP/tSNE)

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")

```
# Finding differentially expressed features (cluster biomarkers)

```{r}
cluster2.markers <- FindMarkers(pbmc, ident.1 = 2, min.pct = 0.25)
head(cluster2.markers, n = 5)
```

```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
```

```{r}
VlnPlot(pbmc, features = c("MS4A1", "CD79A"))

```



```{r}
VlnPlot(pbmc, features = c("NKG7", "PF4"), slot = "counts", log = TRUE)

```

```{r}
FeaturePlot(pbmc, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP",  "CD8A"))
```

```{r}
pbmc.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()
```

Cluster ID	Markers	Cell Type
0	IL7R, CCR7	Naive CD4+ T
1	CD14, LYZ	CD14+ Mono
2	IL7R, S100A4	Memory CD4+
3	MS4A1	B
4	CD8A	CD8+ T
5	FCGR3A, MS4A7	FCGR3A+ Mono
6	GNLY, NKG7	NK
7	FCER1A, CST3	DC
8	PPBP	Platelet

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:10)

DimPlot(pbmc, reduction = "umap")

```

```{r}
new.cluster.ids <- c("Naive CD4+ T-1","Naive CD4+ T-2", "CD14+ Mono", "Memory CD4+ T", "B", "CD8+ T", "FCGR3A+ Mono",  "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

```{r}
pbmc_MemeoryTcell <- subset(pbmc,idents = 'B')

pbmc_Naive2 <- subset(pbmc,idents = 'Naive CD4+ T-2')

```

```{r}
cell_counts <- as.data.frame(table(Idents(pbmc)))
```

#### Нова спроба ####
```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
# Змінити цей параметр resolution
pbmc <- FindClusters(pbmc, resolution = 0.5)
head(Idents(pbmc), 5)

pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")

pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
# дослідити клітинні маркери
pbmc.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)

pbmc.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(pbmc, features = top10$gene) + NoLegend()

# переназвати наші кластери
new.cluster.ids <- c("Naive CD4 T", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
    "NK", "DC", "Platelet")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```


